An Empirical Analysis of the Benefits of Adaptive Gradient Methods

### 2017-12-05 Project/Presentation/main.tex
plotter added
Presentation complete

\section{Our work}
\begin{frame}{Our work}
\begin{itemize}
\item<2->{Implementation of Polyak's heavy ball method(CM), Adagrad, Adadelta, Adam, RMSprop, effectively creating a suite of first-order optimization methods.
          \begin{itemize}
          \item<2->{Implemented in NumPy\footfullcite{numpy}}
          \end{itemize}}
\item<3->{Analyzing behaviour of GD, CM, Adagrad, Adadelta, Adam, RMSprop on selected pathological test functions for optimization}
\item<4->{Experiments on Deep Neural Networks - Multi-layer perceptrons, Autoencoders and Generative Adversarial Networks\footfullcite{gan}}
\end{itemize}
\end{frame}

\subsection{Pathological Functions}
\begin{frame}{Functions which were used}
For our initial experiments, we chose 6 functions\footfullcite{test-funcs}:
\begin{itemize}
\item<1->{Bohachevsky's Functions 1, 2, 3}
\item<2->{Beale's Function}
\item<3->{Rosenbrock's Function}
\item<4->{Styblinsky-Tang's Function}
\end{itemize}
\end{frame}

\begin{frame}{Behaviour with Bohachevsky's Function 1}
\begin{equation}
\label{B1}
f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1}) - 0.4\cos(4\pi x_{2}) + 0.7
\end{equation}
\end{frame}

\begin{frame}{Behaviour with Bohachevsky's Function 2}
\begin{equation}
\label{B2}
f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1})\cos(4\pi x_{2}) + 0.3
\end{equation}
\end{frame}

\begin{frame}{Behaviour with Bohachevsky's Function 3}
\begin{equation}
\label{B3}
f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1} + 4\pi x_{2}) + 0.3
\end{equation}
\end{frame}

\begin{frame}{Behaviour with Beale's Function}
\begin{equation}
\label{BL}
f(\xbold) = (1.5 - x_{1} + x_{1}x_{2})^{2} + (2.25 - x_{1} + x_{1}x_{2}^{2})^{2} + (2.625 - x_{1} + x_{1}x_{2}^{3})^{2}
\end{equation}
\end{frame}

\begin{frame}{Behaviour with Rosenbrock's Function}
\begin{equation}
\label{RB}
f(\xbold;a=1, b=25) = (1 - x_{1})^{2} + 25(x_{2} - x_{1})^{2}
\end{equation}
\end{frame}

\begin{frame}{Behaviour with Styblinsky-Tang's Function}
\begin{equation}
\label{ST}
\displaystyle f(\xbold;n=2) = \sum_{i=1}^{n=2} x_{i}^{2} - 16x_{i}^{2} + 5x_{i}
\end{equation}
\end{frame}
\end{document}

>>>>

\section{Our work}
\begin{frame}{Our work}
\begin{itemize}
\item<2->{Implementation of Polyak's heavy ball method(CM), Adagrad, Adadelta, Adam, RMSprop, effectively creating a suite of first-order optimization methods.
          \begin{itemize}
          \item<2->{Implemented in NumPy\footfullcite{numpy}}
          \end{itemize}}
\item<3->{Analyzing behaviour of GD, CM, Adagrad, Adam and RMSprop on selected pathological test functions for optimization}
\item<4->{Experiments on Deep Neural Networks - Multi-layer perceptrons, Autoencoders and Generative Adversarial Networks\footfullcite{gan}}
\end{itemize}
\end{frame}

\subsection{Pathological Functions}
\begin{frame}{Functions which were used}
For our initial experiments, we chose 5 functions\footfullcite{test-funcs}:
\begin{itemize}
\item<1->{Bohachevsky's Functions 1, 2, 3}
\item<2->{Beale's Function}
\item<3->{Rosenbrock's Function}
\end{itemize}
\end{frame}

\begin{frame}{Behaviour with Bohachevsky's Function 1}
\begin{equation}
\label{B1}
f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1}) - 0.4\cos(4\pi x_{2}) + 0.7
\end{equation}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/B1.eps}
\end{figure}
\end{frame}

\begin{frame}{Behaviour with Bohachevsky's Function 2}
\begin{equation}
\label{B2}
f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1})\cos(4\pi x_{2}) + 0.3
\end{equation}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/B2.eps}
\end{figure}
\end{frame}

\begin{frame}{Behaviour with Bohachevsky's Function 3}
\begin{equation}
\label{B3}
f(\xbold) = x_{1}^{2} + 2x_{2}^{2} - 0.3\cos(3\pi x_{1} + 4\pi x_{2}) + 0.3
\end{equation}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/B3.eps}
\end{figure}
\end{frame}

\begin{frame}{Behaviour with Beale's Function}
\begin{equation}
\label{BL}
f(\xbold) = (1.5 - x_{1} + x_{1}x_{2})^{2} + (2.25 - x_{1} + x_{1}x_{2}^{2})^{2} + (2.625 - x_{1} + x_{1}x_{2}^{3})^{2}
\end{equation}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/BL.eps}
\end{figure}
\end{frame}

\begin{frame}{Behaviour with Rosenbrock's Function}
\begin{equation}
\label{RB}
f(\xbold;a=1, b=25) = (1 - x_{1})^{2} + 25(x_{2} - x_{1})^{2}
\end{equation}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/RB.eps}
\end{figure}
\end{frame}

\subsection{Deep Learning Experiments: MLP Classifiers}
\begin{frame}{Multi-Layer Perceptron Classifiers}
\begin{itemize}
\item<1->{Took Multilayer perceptron classifiers for the MNIST\footnote{\url{http://yann.lecun.com/exdb/mnist/}} and CIFAR10\footnote{\url{https://www.cs.toronto.edu/~kriz/cifar.html}} datasets}
\item<2->{Achieved \(\approx 95\%\) on MNIST and \(\approx 60\%\) on CIFAR10
          \begin{itemize}
          \item<3->{Not State-of-the-art, only to analyze the behaviour of the optimization methods}
          \end{itemize}
         }
\item<3->{Activation used between the layers was \texttt{ReLU}}
\item<4->{Trained for 10000 iterations, chose optimal learning rate from \(\{0.5, 0.2, 0.1, 0.05, 0.02, 0.01, 0.005, 0.002, 0.001, 0.0005, 0.0002, 0.0001\}\)}
\end{itemize}
\end{frame}

\begin{frame}{MNIST with Architecture: \texttt{784-500-500-500-10} - Training Loss vs Mini-batches}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/mnist-1.pdf}
\end{figure}
\end{frame}

\begin{frame}{MNIST with Architecture: \texttt{784-1000-1000-1000-10} - Training Loss vs Mini-batches}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/mnist-2.pdf}
\end{figure}
\end{frame}

\begin{frame}{CIFAR10 with Architecture: \texttt{3072-1000-500-10} - Training Loss vs Mini-batches}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/cifar10-1.pdf}
\end{figure}
\end{frame}

\begin{frame}{CIFAR10 with Architecture: \texttt{3072-1000-1000-1000-10} - Training Loss vs Mini-batches}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/cifar10-2.pdf}
\end{figure}
\end{frame}

\subsection{Deep Learning Experiments: Autoencoders}
\begin{frame}{Deep Autoencoders}
\begin{itemize}
\item<1->{Used a multilayer perceptron deep autoencoder for encoding MNIST images}
\item<2->{Activation used between the layers was \texttt{sigmoid}}
\item<3->{Trained for 10000 iterations, chose optimal learning rate from \(\{0.5, 0.2, 0.1, 0.05, 0.02, 0.01, 0.005, 0.002, 0.001, 0.0005, 0.0002, 0.0001\}\)}
\end{itemize}
\end{frame}

\begin{frame}{Autoencoder architecture: \texttt{784-1000-500-1000-10} - Training Loss vs Mini-batches}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/auto-1.pdf}
\end{figure}
\end{frame}
\begin{frame}{Autoencoder architecture: \texttt{784-1000-500-250-30-30-250-500-1000-10}\footfullcite{hinton} - Training Loss vs Mini-batches} % add citation
\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{./images/auto-2.pdf}
\end{figure}
\end{frame}

\subsection{Deep Learning Experiments: GANs}
\begin{frame}{Deep Convolutional Generative Adversarial Networks\footfullcite{dcgan}} %add citation
\begin{itemize}
\item<1->{Used a specialized architecture for generator and discriminator for MNIST digits based on Deep Convolutional GANs}
\item<2->{Run for 50000 iterations, and noted the variations of generator loss and discriminator loss with time for Adam optimizer and Nesterov's Accelerated Gradient}
\end{itemize}
\end{frame}

\begin{frame}{Generator Loss with time}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/genr-gan.pdf}
\end{figure}
\end{frame}
\begin{frame}{Discriminator Loss with time}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{./images/disc-gan.pdf}
\end{figure}
\end{frame}

\begin{frame}{Quality of Images Generated}
\begin{columns}
\column{0.475\linewidth}
\centering
\textbf{Using ADAM optimizer}
\includegraphics[width=0.75\textwidth]{./images/adam-dcgan.png}
\column{0.475\linewidth}
\centering
\textbf{Using Accelerated Gradient}
\includegraphics[width=0.75\textwidth]{./images/nag-dcgan.png}
\end{columns}
\end{frame}

\section{Conclusion}
\begin{frame}{Our conclusions from the experiments conducted}
\begin{itemize}
\item<1->{Experiments show that adaptive gradient methods in fact, only have a marginal benefit}
\item<2->{This concurs with the claim put forward by Wilson \textit{et al}\footfullcite{marginaladaptivegradients} about the marginal benefits of Adaptive Gradient Methods}
\item<3->{Hype or Hallelujah? \emph{Hype} it is!!}
\end{itemize}
\end{frame}

\begin{frame}
\Huge{Thank You\footnote{Code Available at \url{https://github.com/vishwakftw/CS6230-OML}}}
\end{frame}
\end{document}
